---
title: "Lab 09: Logistic Regression"
subtitle: ""
author: ""
date: ""
editor_options: 
  chunk_output_type: console
output:
  pdf_document: default
---

```{r set_up, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

## Packages

```{r packages}
library(tidyverse)
library(broom)
```

## Data

```{r data}
retirement <- read_csv("data/retirement_data.csv")
```

The 2018 election saw a near-record number of [retirements](https://www.pewresearch.org/fact-tank/2018/04/11/near-record-number-of-house-members-not-seeking-re-election-in-2018/) 
from Congress. In this lab, you will be working with data from that election.

The variables in this dataset are:

- `stcd`: an ID number for the congressional district. Districts are listed in 
          alphabetic order (e.g., 101 is Alabama's first district).

- `dpres`: the two-party percentage of vote received by Hillary Clinton 
           in the district in 2016. (Third party vote has been eliminated here, 
           this represents Democratic vote/ (Democratic + Republican Vote).)

- `retiring`: a dummy variable indicating whether the representative retired in 
              2018. The includes both representatives who left politics 
              altogether and those who sought higher office, as well as some 
              who resigned early.

- `gopseat`: a dummy variable indicating whether the seat was held by a 
             Republican.

- `age`: the age of the representative at the end of 2018.

# Exercises

## Exercise 1

```{r ex-1}
retirement <- retirement %>%
  filter(!is.na(retiring))

retirement <- retirement %>%
  mutate(presvote = ifelse(gopseat == 1, 100 - dpres, dpres))

retirement_mod <- glm(retiring ~ gopseat + age + presvote,
                      data = retirement,
                      family = "binomial")
tidy(retirement_mod)
```

$$\text{log}\left(\frac{\hat{p}}{1-\hat{p}}\right) = 0.023 + 0.051~gopseat + 0.0013~age - 0.049~presvote$$

## Exercise 2

```{r ex-2}
tidy(retirement_mod) %>%
  mutate(estimate= exp(estimate))
```

Holding all else constant, for Republicans, the odds of retirement is 
$e^{0.507} = 1.660$ times the odds of retirement for Democrats. 

Holding all else constant, for every one year increase in age, the odds of 
retirement are expected to be multiplied by $e^{0.013} = 1.013$.

Holding all else constant, for every one percentage point increase in 
presidential vote, the odds of retirement are expected to be multiplied by 
$e^{-0.487} = 0.952$.

## Exercise 3

```{r ex-3}
mean_age <- mean(retirement$age)

tibble(
  age      = rep(mean_age, times = 126),
  gopseat  = rep(c(0, 1), each = 63),
  presvote = rep(33:95, times = 2)
) %>%
  augment(retirement_mod, newdata = .) %>%
  mutate(p = exp(.fitted) / (1 + exp(.fitted))) %>% 
  ggplot(aes(x = presvote, y = p, color = factor(gopseat))) +
  geom_line() +
  scale_color_manual(values = c("#0015BC", "#DE0100")) +
  labs(title = "Probability of a Retirement",
       y = "Predicted Probability of Retirement", x= "Presidential Vote",
       color = "GOP Seat") +
  theme_bw()
```

For all levels of presidential vote, the probability of a Republican retirement
is higher than that of a Democratic retirement, although this gap is especially
large for marginal seats (i.e., where the values of presidential vote are low). 
As presidential vote increases, the probability of a retirement decreases. 

## Exercise 4

```{r ex-4}
tibble(
  age      = 78,
  gopseat  = 0,
  presvote = 70.75
  ) %>%
  augment(retirement_mod, newdata = .) %>%
  mutate(p = exp(.fitted) / (1 + exp(.fitted))) %>%
  pull("p")
```

According to our model, Price has a 8.25% predicted probability of retirement 
in 2018.

## Exercise 5

```{r ex-5}
retirement_mod %>%
  augment(data = retirement) %>%
  mutate(prob = (exp(.fitted) / (1 + exp(.fitted)))) %>%
  select(stcd, prob, retiring) %>%
  arrange(desc(prob)) %>% 
  slice(1)
```

No representative is given over a 50% probability of retirement. The district 
with the code 3805 (this is PA's 5th district, no need to give answer in words) 
was given the highest probability of retirement and they did indeed retire. 
