---
title: "Lab 02"
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---

# Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
``` 

# Exercise 1

```{r ex-1}
ggplot(mpg, aes(x = hwy, y = cty, color = class)) + 
  geom_point() +
  labs(title = "City versus highway miles per gallon",
       subtitle = "1999 - 2008",
       x = "Highway miles per gallon", y = "City miles per gallon",
       color = "Car class")
```

# Exercise 2

```{r ex-2}
ggplot(mpg, aes(x = hwy, y = cty, color = class)) + 
  geom_jitter() +   
  labs(title = "City versus highway miles per gallon",
       subtitle = "1999 - 2008",
       x = "Highway miles per gallon", y = "City miles per gallon",
       color = "Car class")


#ggplot(mpg, aes(x = hwy, y = cty, color = class)) + 
#  geom_point(position = "jitter") +   
#  labs(title = "City versus highway miles per gallon",
#       x = "Highway miles per gallon", y = "City miles per gallon")
```

Some cars in this dataset have the same highway and city miles per gallon, so 
the points are on top of each other. For example, the Toyota Camry Solara and
the Volkswagen Passat both have 18 city and 26 highway miles per gallon, but 
there is only one point on the scatterplot. We can't see the trend because some
of the points are hidden. 

Jittering solves this by adding a small amount of random variation to each
point. The scatterplot no longer perfectly represents the data, but we get a
much better sense of the trend.

The first scatterplot is dishonest, as it is not displaying all of the 
relevant data to a reader.

# Exercise 3

```{r ex-3}
ggplot(mpg, aes(x = hwy, y = cty)) + 
  geom_point() + 
  facet_wrap( ~ drv) +
    labs(title = "City versus highway miles per gallon",
         subtitle = "faceted by drive train",
       x = "Highway miles per gallon", y = "City miles per gallon")
```

Quite a few things to notice here! 

Higher highway miles per gallon is associated with higher city miles per gallon
for all types of drive train.

Cars with front wheel drive have more variation in both city and highway miles 
per gallon than four wheel or rear wheel drive cars. They also have generally
the highest miles per gallon.

Add more.

# Exercise 4

```{r ex-4}
ggplot(mpg, aes(x = class, y = cty)) +
  geom_boxplot() +
    labs(title = "Distribution of city miles per gallon by class",
       x = "Class", y = "City miles per gallon")
```

We observe about what we expect based on car knowledge. Heavier cars (pickups
and sports utility vehicles) have a lower distribution of city miles per gallon
and smaller cars (subcompacts, compacts, and midsize cars) have the highest
distribution of city miles per gallon. Minivans and two-seaters are in the
middle. 

To better see the relationship, we can orient the axis in a more sensible way
than alphabetically. Here we arrange the cars in order of descending median
city miles per gallon.

```{r ex-4-axes}
ggplot(mpg, aes(x = fct_reorder(class, cty, median, .desc = TRUE), y = cty)) +
  geom_boxplot() +
  labs(title = "Distribution of city miles per gallon by class",
       x = "Class", y = "City miles per gallon")
```

# Exercise 5

```{r ex-5}
ggplot(mpg, aes(x = class, fill = drv)) + 
  geom_bar(position = "fill") +
  labs(title = "Segemented bar plots of car class",
       subtitle = "filled by drive train",
  x = "Class", y = "Count", fill = "Drive Train")
```

All two-seaters have rear wheel drive. A higher proportion of compacts, 
midsize cars, and minivans have four wheel drive, and pickups and sports utility
vehicles tend to be all-wheel drive. Subcompacts are the only car class with
all three types of drive train.

# Exercise 6

```{r ex-6}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), size = 0.50) + 
  facet_wrap(~ class, nrow = 2) +
  theme_bw() +
  labs(title = "Highway mpg versus displacement",
       subtitle = "faceted by class",
         x = "Displacement (litres)", y = "Highway mpg")
```