---
title: "Lab 04 Solutions"
author: ""
date: ""
editor_options: 
  chunk_output_type: console
output:
  pdf_document: default
---

# Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(sf)
``` 

# Data

```{r load-data, message = FALSE}
fl_votes <- st_read("data/fl_votes.shp", quiet = TRUE)

fl_votes %>%
  slice(1:6)
```

# Exercise 1

The `fl_votes` data is not tidy. The year of the election (2016 and 2020) is
embedded in the columns `rep16`, `dem16`, `rep20`, and `dem20`. We also have
a new `geometry` column (actually a list-column) with the simple feature 
geometries for each Florida county.

# Exercise 2

```{r add-winner}
fl_votes <- fl_votes %>%
  mutate(winner20 = if_else(rep20 > dem20, "Donald Trump", "Joe Biden"))
```

We probably should use `case_when()` here to be safe and handle situations
with a tie.

# Exercise 3

```{r fl-plot-1}
ggplot(fl_votes) +
  geom_sf(aes(fill = winner20)) + 
  scale_fill_manual(values=c("#DE0100", "#0015BC")) +
  labs(title = "2020 Presidential election results",
       subtitle = "Florida counties",
       fill = "Winner") + 
  theme_bw()
```

# Exercise 4

```{r fl-props}
fl_votes <- fl_votes %>%
  mutate(prop_rep16 = rep16 / (rep16 + dem16), 
         prop_rep20 = rep20 / (rep20 + dem20))
```

# Exercise 5

```{r fl-plot-2}
ggplot(fl_votes) + 
  geom_sf(aes(fill = prop_rep20)) + 
  scale_fill_gradient(low = "#0015BC", high = "#DE0100") +
    labs(title = "2020 Presidential election results",
       subtitle = "Florida counties",
       fill = "Republican share of vote") + 
    theme_bw()
```

# Exercise 6

```{r fl-change}
fl_votes <- fl_votes %>%
  mutate(diff_rep = prop_rep20 - prop_rep16)
```

# Exercise 7

```{r fl-plot-3}
ggplot(fl_votes) +
  geom_sf(aes(fill = diff_rep)) + 
  scale_fill_gradient2(low = "#0015BC", midpoint = 0, high = "#DE0100") +
    labs(title = "Change in Republican share of two party vote",
       subtitle = "between 2016 and 2020",
       fill = "Change in R share of vote") + 
    theme_bw()
```

# Exercise 8

There are many possibilities here. Some things to notice are below (there are 
more).

- Using two colors in the first plot is misleading - not everyone in a county
  voted for a single candidate.
  
- Second plot is an improvement over the first. We see a more purple state 
  compared to the first plot. Cities (mostly Miami, Tallahassee, Tampa, Orlando, 
  and Jacksonville) favor Clinton, while more rural areas favor Trump,
  particularly in the Northern part of the state. 
  
- Second plot is still misleading since counties are sized according to area
  rather than population - resizing counties, adding vertical bars corresponding
  to the number of votes cast, or something similar would be helpful. The 
  election was more close than this map indicates (51.6% Trump, 48.3% Biden 
  considering only the two party vote).
  
- There is a lot going on in the third plot. Moderate shifts towards the 
  Democratic candidate occurred in the panhandle, Jacksonville, Fort Myers, 
  and the Eastern coast of the state, but Orlando and Miami shifted towards the
  Republican candidate, likely due to the high population of Cuban and 
  Venezuelan voters in these areas. 
  
- We note that the third plot indicates **change** when comparing the share of
  the two party vote between 2016 and 2020. Bright red Miami-Dade county is a
  heavily Democratic county, but the Republican share of the two-party vote 
  increased from 34.8% to 46.3%. Similarly, Escambia county in the far West
  of the state is still heavily Republican, but the Republican share of the
  two party vote decreased from 60.7% to 57.6%.
