---
title: "Lab 03"
editor_options: 
  chunk_output_type: console
output:
  pdf_document: default
---

# Load packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
``` 

# Data

```{r load-data, message = FALSE}
lego <- read_csv("data/lego.csv")
```

# Exercise 1

```{r ex-1}
lego <- lego %>%
  filter(retail_price > 0, 
         pieces > 0, 
         !is.na(retail_price), 
         !is.na(pieces),
         !is.na(year))

lego
```

Should have a careful consideration of the implications of removing rows 
containing missing data.

Sets with missing price or piece information may differ from sets without 
missing information. As one possibility, older sets may be more
likely to not have retail price information recorded. If this is the case, 
removing sets with missing information leads to biased results.

# Exercise 2

```{r ex-2}
lego %>% 
  arrange(desc(retail_price)) %>% 
  slice(1:3) %>%
  print(width = Inf)
```

The three most expensive lego sets are the Millennium Falcon (\$800 with 7,541
pieces), the Connections Kit (\$755 with 2,455 pieces), and the Death Star 
(\$500 with 4,016 pieces).

# Exercise 3

```{r ex-3}
lego <- lego %>%
  mutate(price_per_piece = retail_price / pieces)
```

# Exercise 4

```{r ex-4}
lego %>%
  arrange(desc(price_per_piece)) %>%
  slice(1:5) %>% 
  select(name, themegroup, theme, pieces, price_per_piece)
```

These sets are electronic components in either the Technical or Educational 
themegroup: computers, programming bricks, batteries, etc. They contain only 
a single piece but are expensive.

# Exercise 5

```{r ex-5}
lego %>%
  filter(theme == "The Lord of the Rings") %>%
  group_by(subtheme) %>%
  arrange(retail_price) %>%
  slice(1, n()) %>%
  select(name, subtheme, retail_price, pieces)
```

# Exercise 6

```{r ex-6}
yearly_trends <- lego %>%
  group_by(year) %>%
  summarize(median_price = median(price_per_piece, na.rm = TRUE),
            count = n()) 
```

# Exercise 7

```{r ex-7}
yearly_trends %>%
  ggplot(aes(x = year, y = median_price, size = count)) + 
  geom_point(alpha = 0.50) +
  geom_line(size = 0.50, alpha = 0.50) + 
  labs(x = "Year", y = "Median Price ($)", size = "Count", 
       title = "Change in price per piece of LEGO sets over time")
```

The median price per piece has remained roughly stable over time since about 
1990 at about ten cents per piece. There are a few years in the 60's, 70's, and 
80's with a small sample size (only a few sets produced) and a high median price
per piece.