---
title: "Lab 06 Solutions and Rubric"
subtitle: ""
author: ""
date: ""
editor_options: 
  chunk_output_type: console
output:
  pdf_document: default
---

```{r set_up, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
set.seed(352790)
```

# Packages

```{r load_packages}
library(tidyverse)
library(infer)
``` 

# Data

In this lab, you'll work with a couple of datasets.

The `ToothGrowth` dataset can be loaded into R with `data("ToothGrowth)`. It
contains data on the length of odontoblasts (cells responsible for tooth growth) 
in 60 guinea pigs. Each animal received one of three dose levels of vitamin C 
(0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice (`OJ`) or 
ascorbic acid (`VC`) (a form of vitamin C and coded as VC). For the purposes of 
this lab, we will ignore the `dose` variable.

```{r load_tooth_data}
data("ToothGrowth")
```

The second dataset is a subset of `gss_cat` from the `forcats` package. It
contains categorical variables from the General Social Survey in 2014.

```{r load_gss_data, echo=TRUE}
gss_2014 <- gss_cat %>% 
  filter(year == 2014)
```

# Exercises

## Exercise 1 

- **Population**: Guinea pigs that receive vitamin C
- **Parameter**: population mean, $\mu$, length of odontoblasts
- **Sample**: Guinea pigs assumed to be randomly selected as given in `ToothGrowth`

The observed **sample mean** and **sample size** are given below.

```{r exericse_1}
ToothGrowth %>% 
  summarise(mean_length = mean(len),
            sample_size = n())
```


## Exercise 2 

```{r exercise_2}
ToothGrowth %>% 
  specify(response = len) %>% 
  generate(reps = 10000) %>% 
  calculate(stat = "mean") %>% 
  get_ci(level = 0.99)
```

We are 99% confident that the mean length of odontoblasts for guinea pigs that
receive vitamin C is captured by the interval (16.4, 21.4).

## Exercise 3 

```{r exercise_3_obs}
ToothGrowth %>% 
  specify(len ~ supp) %>% 
  calculate(stat = "diff in means", order = c("OJ", "VC"))
```

```{r exercise_3_viz}
boot_diff_means <- ToothGrowth %>% 
  specify(len ~ supp) %>% 
  generate(reps = 10000) %>% 
  calculate(stat = "diff in means", order = c("OJ", "VC"))

visualise(boot_diff_means) +
  labs(x = "Difference in means (OJ - VC)", y = "Count")
```

```{r exercise_4_ci}
get_ci(boot_diff_means, level = 0.95)
```

We are 95% confident that mean difference in odontoblast length for guinea pigs
given OJ versus ascorbic acid is captured by the interval (-0.125, 7.34).

## Exercise 4 

Let $\alpha = 0.01$ for the hypothesis test of

$$H_0: p = 0.48$$
$$H_A: p \ne 0.48$$

```{r exercise_5_observed}
p_hat <- gss_2014 %>% 
  mutate(marital = ifelse(marital == "Married", "Married", "Not married")) %>%
  count(marital) %>% 
  mutate(proportion = n / sum(n)) %>% 
  filter(marital == "Married") %>% 
  select(proportion)
```

```{r exercise_5_null}
gss_2014 %>% 
  mutate(marital = ifelse(marital == "Married", "Married", "Not married")) %>% 
  specify(response = marital, success = "Married") %>% 
  hypothesise(null = "point", p = 0.48) %>% 
  generate(reps = 1000) %>% 
  calculate(stat = "prop") %>% 
  get_p_value(obs_stat = p_hat, direction = "two-sided")
```

Given that the p-value is greater than the significance level, we fail to
reject the null hypothesis. That is, we fail to reject the claim that the
proportion of U.S. adults who are married is 0.48 at the 0.01 significance 
level.

## Exercise 6 

Since the data led us to fail to reject the null hypothesis, the only type of
"error" that could have been made is a Type II error.

## Exercise 7 

Yes. If we require weaker evidence to reject the null hypothesis, then given 
that the p-value was 0.016, we would reject $H_0$ at the 0.10 significance
level.

# References

"Infer - Tidy Statistical Inference". Infer.Netlify.App, 2021, 
https://infer.netlify.app/index.html.