---
title: "Lab 08: Linear Regression"
subtitle: ""
author: ""
date: ""
editor_options: 
  chunk_output_type: console
output:
  pdf_document: default
---

```{r set_up, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

# Packages

```{r packages}
library(tidyverse)
library(broom)
```

# Exercises

## Exercise 1 

```{r ex-1}
ggplot(data = midwest, aes(x = percollege , y = percbelowpoverty)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = " College Education and Poverty Rate in Midwestern Counties",
       x = "Percentage with a College Degree", y = "Poverty Rate")
```

## Exercise 2

```{r ex-2}
college_model <- lm(percbelowpoverty ~ percollege, data = midwest)
tidy(college_model)
```

$$\widehat{PercentBelowPoverty} = 16.7 - 0.23~CollegePercentage$$

Slope: For each additional percentage point of people who attended college in a
Midwest county, the poverty rate is expected to be lower, on average by 0.23
percentage points. 

Intercept: Counties with 0% of people with college degrees are expected, on 
average, to have 16.7% of people below poverty.

## Exercise 3

```{r ex-3}
glance(college_model) %>% 
  select(r.squared)
```

Roughly 8% of the variability in the percentage of people below poverty can be
explained by the percentage of people with a college degree. This is a low $R^2$
value.

## Exercise 4

```{r ex_4}
tidy(college_model, conf.int = TRUE, conf.level = 0.95) %>%
  filter(term == "percollege") %>% 
  select(starts_with("conf"))
```

We are 95% confident that for every additional percentage point of people in a
Midwest county who have a college degree, the percentage of people below poverty
is expected to decrease, on average, between 0.157 and 0.306 percentage points.

## Exercise 5

```{r ex_5}
college_model %>%
  augment(data = midwest) %>%
  filter(county == "SUMMIT") %>%
  select(percbelowpoverty, .fitted, .resid)
```

The model predicts that 11% of people in Summit County, Ohio will be below 
poverty, but the real value here is 12.1%, a difference of 1.16 percentage 
points. The difference here is the residual.

## Exercise 6

```{r ex-6}
state_model <- lm(percbelowpoverty ~ factor(state), data = midwest)
tidy(state_model)
```

Slopes: These coefficients describe the expected difference in poverty in that 
state compared to the baseline level. Here the baseline state is Illinois, so
predicted poverty, on average, in an Indiana county would be 2.76 percentage 
points lower than in Illinois. On average, it would be 1.15 percentage points
higher in Michigan, 0.04 percentage points lower in Ohio, and 1.18 percentage
points lower in Wisconsin. Only the Indiana coefficient is statistically
significant.  

Intercept: A county located in Illinois is expected, on average, to have 13.08
percent of people below poverty. This is significant at the 0.05 level.

$$\widehat{PercentBelowPoverty} = 13.1 - 2.76~State_{IN} + 1.15~State_{MI} -0.04State_{OH}-1.18~State_{WI}$$

# References

Midwest Demographics. Dataset in `ggplot2`. https://ggplot2.tidyverse.org/reference/midwest.html